---
title: "Homework-16"
author: "ML"
format: html
editor: visual
---

10/09/2024

\## Homework 16

Reinforcing Chapter 3 Concepts

set things up

```{r}
rm(list = ls()) #clears environment
# establish packages
library (tidyverse)
library (here)
```

Create a new dataframe called "cover" by importing the csv file called “donnerville_percent_cover.csv”

```{r}
cover <- read.csv (here("data", "donnerville_percent_cover.csv")) 
```

examine the data set using at least 2 functions

```{r}
#1
summary (cover) # median, mean, min, max, and interquartile ranges and sample sizes
#2 
names (cover) # column names
```

Make Site, Plot type and cover_type into factor variables.

```{r}
cover$Site <- as.factor (cover$Site)
cover$Plot.type <- as.factor (cover$Plot.type)
cover$cover_type <- as.factor (cover$cover_type)
```

Make Date into a date variable.

```{r}
?as.Date() # i forgot how this works...let's check
# cover$Date <- as.Date(cover$Date,"%d%b%Y")
# this turned the values to NA, let's restart

# i'm going to try using which() to fix the data and try this method again
fixTreatment <- which (cover$Date == ("13-Sep-17")) 
  cover$Date [fixTreatment] <- "13jan2017"
  
# now run the Date code again
cover$Date <- as.Date(cover$Date,"%d%b%Y") # much better
```

Do some error checking on your data and fix any mistakes you find.

```{r}
# definitely an issue with a Site name 
levels (cover$Site)

fixTreatment1 <- which (cover$Site %in% c("DONNERVILLE - 1" , "donnerville 1")) 
  cover$Site [fixTreatment1] <- "Donnerville 1"

fixTreatment2 <- which (cover$Site %in% c("Donnerville 2" , "donnerville 2")) 
  cover$Site [fixTreatment2] <- "Donnerville 2"  
  
fixTreatment3 <- which (cover$Site %in% c("donnerville-3" , "Donnerville 3")) 
  cover$Site [fixTreatment3] <- "Donnerville 3"
  
cover$Site <- factor (cover$Site)
levels(cover$Site)

# and cover_type
levels (cover$cover_type)

fixTreatment4 <- which (cover$cover_type %in% c("bare" , "Bare")) 
  cover$cover_type [fixTreatment4] <- "Bare"
fixTreatment5 <- which (cover$cover_type %in% c("bryophtes" , "Bryophytes")) 
  cover$cover_type [fixTreatment5] <- "Bryophytes"
fixTreatment6 <- which (cover$cover_type == ("vasc ")) #GRRRR there is an extra space here that was hard to find
  cover$cover_type [fixTreatment6] <- "Vascular"
  
cover$cover_type <- factor (cover$cover_type)
levels(cover$cover_type)

```

Determine the average percent cover by cover type (bare, bryophytes, vascular plants) for Donnerville 1.

```{r}
# okay i want to find the mean percent for each cover_type (a column) for Donnerville 1 (a value in column Site)
cover %>% # work with cover data
  group_by (cover_type)%>% # group data into cover types
    filter (Site=="Donnerville 1") %>% # get all the rows where Site = Donnerville 1
      summarise (
          meancover = mean (Percent)) # find the means of these data
```

Determine the average percent vascular plant cover across all sites.

```{r}
# finding the mean percent of vascular (a value in cover_type) 
cover %>% # work with cover data
    filter (cover_type=="Vascular") %>% # get all the rows where covertype = vascular
      summarise (
          meancover2 = mean (Percent)) # find the means of these data
```
